\chapter{Everything I Believe Is False}

\lettrine{O}{ddly} enough, it might have been easier explaining to his
dad that an owl had grabbed the letter after all.

"What? \emph{Mrs.~Figg?}" Professor Verres-Evans's shock was
spectacular. Harry empathized completely. He sat at the
table between the living room and kitchen, somewhat in a daze.

"Did she say what time they would be coming?" Petunia
asked. She checked the pot roast and ran her hands over
her hair, as if expecting someone to ring the doorbell at any minute.

"We've known Mrs.~Figg for ten years," Harry's dad said.
"She's a perfectly reasonable woman. Why on earth would
she—"

"No, Mum, she just said they'd be here in a `jiffy or two.'
I don't know how far they're coming from, but it's probably not
going to be{\el}" Harry trailed off as he realized that, given
the hypothesis being tested, it might well be before the pot
roast was done. He tried to remind himself how silly that was—
teleportation would break so many laws of physics there
might as well not be any—but ever since he'd heard Mrs.~Figg
speak the name "Hogwarts," his brain
didn't seem to be working properly. A part of his mind
took note of his dad's comment about knowing Mrs.~Figg
for ten years. Had she moved here the year Harry had
been adopted? That seemed significant, if he could only
wrestle his mind into considering how.

"Maybe \emph{she} sent the letters," Dad said. He began to to
pace the limited floorspace of the living room, stepping
around books with the unconscious ease of memory. "Or perhaps
she's part of the same cult your sister was in—"

"Well, we'd better set an extra place just in case." Mum put a
stack of plates in front of Harry, and he set the table for
four, placing each fork and knife with an inordinate amount
of attention. Dad's theories made sense, of course, more
sense than his did, but his strange certainty continued to
color all his thoughts as he set out the cups.

A knock at the door froze them all in place.

Professor Verres-Evans was the first to thaw. He straightened, squared his
shoulders, and walked to the front door, dignity fully
reinforced by his casual tweed outfit.

Mum wiped her hands on a towel and followed. Harry
rushed after them, wondering if it would be Mrs.~Figg yet
knowing somehow that it wasn't. Dad peered through the
peephole and recoiled as if stung. Harry's anticipation redoubled.

"Yes, who's there?" Professor Verres-Evans's voice did not tremble.

"Professor Minerva McGonagall," said a formal, Scottish
voice, and Michael twitched. Harry wondered why, until his
dad opened the door.

Professor McGonagall was an older woman, perhaps in her
sixties, with greying hair in a severe bun and square
spectacles perched on her nose. She looked every inch
the professor she claimed to be, but for two things: she
wore a black robe of some rich fabric, and her hat was
decidedly pointy.

Harry grinned. His father's mental image of "professor" had
just been severely abused.

"Come in, please," Petunia said with a smile. "Supper's
almost ready, if you're hungry."

"I ate, thank you." Professor McGonagall said as she walked
inside. Harry and his parents stepped back to let her through.

"I'm Petunia Evans-Verres, so nice to meet you{\el}" The two
walked down the hall, leaving Harry and his dad by the
door. Harry closed it, then exchanged a look with his father.

"What do you reckon?" Harry whispered. "Time to call the white coats?"

Dad snorted and clapped him on the shoulder with a grin.
"Come on, let's get this foolishness done with." They
followed the women into the living room.

"Got an experiment in mind?" Harry asked. He was still
feeling off balance. That strange certainty was stronger
now, as if all this were a formality, and he had already
accepted that the woman in their house was a witch,
without quite being able to grasp what that would mean in
a practical sense.

"I can't imagine anything I would come up with that she
wouldn't make an excuse for," his father said, still keeping
his voice low. Harry nodded in response, and then he decided to be forthright
with their guest, who stood expectantly in the living room.
She eyed the multitude of books with what looked to be
an admiring air, which Harry found reassuring.

"Good evening, Professor McGonagall. As I'm sure you're
aware—" Harry stopped. He actually wasn't sure \emph{what} she
was aware of. Had she received his letter? What had Mrs.~Figg done,
read it to her over the phone? Perhaps she
had stopped next door to retrieve it before coming here{\el}
Harry stifled his questions and began again. "I'm Harry
Potter-Evans-Verres. We were surprised by your letter, and
have some doubts about its validity. Mum says she's seen
magic before, but neither Dad nor myself have. If you
could demonstrate the quality of your magic to us, that
would be a good first step."

Professor McGonagall was watching Harry with an amused
expression as he spoke. "Of course. I would be happy to."
She pulled a thin wooden stick out of her sleeve with
practiced grace, and Harry blinked. He hadn't seen the
shape of it against the material, and it should have fallen
out if it wasn't held there somehow. "Is there something
specific that would persuade you?"

Still preoccupied with her sleight of hand, it took Harry a
second to realize she was holding a \emph{magic wand}, and he
said the first thing that popped into his head: "Can you
shoot fire out of that?"

"Harry!" Mum said with some alarm, and Professor
McGonagall's lips twitched in a brief smile.

"I could, but I think that would be dangerous." She
glanced pointedly at their surroundings. "How about
something less destructive?"

"Of course," Harry said, cheeks red. "Er{\el} did you fly
here? I didn't hear a car, and unless you live nearby I
can't imagine how else you got here so quickly. If you
could just{\el} hover a bit? That might help. Wait, on second
thought, levitate Dad."

Professor Verres-Evans gave Harry an approving nod, and
stepped forward to face their guest with his arms crossed.
Professor McGonagall lifted her wand, and Harry realized
his mistake. "Wait!" he said. She lowered her wand, raising
an eyebrow. "I want to make sure we do this right." He
thought about it for a second while everyone watched him.

"Now, just to be clear," Harry said to his father. "If Professor
McGonagall does levitate you, when you know you haven't
been attached to any wires, that's going to be sufficient
evidence. You're not going to turn around and say that it's
a magician's trick. That wouldn't be fair play. If you feel
that way, you should say so now, and we can ask her to
do something else instead."

Dad nodded, smiling good-naturedly. "Agreed."

"And you, Mum, your theory says that the professor
should be able to do this, and if that doesn't happen,
you'll admit you're mistaken. Nothing about how magic
doesn't work when people are skeptical of it, or anything
like that."

Mum glanced at Professor McGonagall's wand and nodded.

"Is that sufficient, Mr. Potter?" Professor McGonagall said.
"Shall I go ahead and demonstrate?"

"Sufficient? Probably not, but it should do for now," Harry
said. Once he saw her methodology he could better decide
how to isolate her actions and their relation to the result{\el}
assuming there was a result. Did he really expect his
father to start floating? "Proceed, please."

"Is there anything you'd like me to do?" Professor
Verres-Evans said, still smiling. "Think light thoughts,
perhaps?"

"No need, thank you," Professor McGonagall replied, and
then, "\emph{Wingardium Leviosa}."

There was a silent pause that Harry knew he would
always remember{\el} the moment when his world utterly
changed. Everything was still, as if suspended in crystal:
he and his mother staring in shock, the witch holding her
wand pointed up at his father, who hung a respectable
three feet off the ground in complete defiance of gravity.

Harry looked up at his father.

"Huh."

His father looked down at him.

And then Professor Verres-Evans looked back at Professor
McGonagall and said, in a voice Harry had never heard him
use, "All right, you can put me down now, thank you."

Harry's father was lowered carefully to the ground, and the
moment was ended. The universe continued on just as it had before.

Harry ruffled a hand through his dark hair. Maybe it was
just that strange part of him which had \emph{already} been
convinced, but{\el} "That's a bit of an anticlimax," Harry said.
"You'd think there would be some kind of more dramatic
mental event associated with updating on an observation of
infinitesimal probability—" Harry stopped himself. Mum and
the witch were looking at him oddly. Dad slowly sat down,
not even bothering to move the book from the chair as
he stared at the piece of wood in Professor McGonagall's
hand. "I mean, with finding out that everything I believe is false."

Seriously, it should have been more dramatic. His brain
ought to have been flushing its entire current stock of
hypotheses about the universe, none of which allowed this
to happen. But instead it just seemed to be going, \emph{All
right, I saw the Hogwarts Professor wave her wand and
make Dad rise into the air, now what?}

The witch was smiling benevolently upon them, looking
quite amused. "Would you like a further demonstration, Mr.~Potter?"

"You don't have to," Harry said. "We've performed a definitive
experiment. That wasn't some trick with mirrors, it wasn't
hypnotic suggestion, he actually lifted off the ground, we all
saw it{\el} but{\el}" Harry hesitated. He couldn't help himself.
Actually, under the circumstances, he \emph{shouldn't} be helping
himself. It was right and proper to be curious. "What else
\emph{can} you do?"

"Besides shoot fire, you mean?"

Dad looked as alarmed as Mum had a moment ago.

"Yes, besides that." Though Harry actually would love to
see it. He was starting to get excited. Would he really be
able to fly? Conjure fire at will? How? Did he accelerate
the atoms in the air until they combusted? \emph{Maybe I use
my body heat to—}

Professor McGonagall turned into a cat.

Harry scrambled back unthinkingly, backpedaling so fast that he tripped over a
stray stack of books and landed hard on his bottom. His
hands came down to catch himself without quite reaching properly, and there was
a warning twinge in his shoulder as the weight came down unbraced.

At once the small tabby cat morphed back up into a robed woman. "I'm sorry,
Mr.~Potter," said the witch, sounding sincere, though the corners of her lips
were twitching upwards. "I should have warned you."

Harry was breathing in short gasps. His voice came out choked.
"\emph{You can't DO that!}"

"It's only a Transfiguration," said Professor McGonagall. "An Animagus
transformation, to be exact."

"You turned into a cat! A \emph{SMALL} cat! You violated Conservation of
Energy! That's not just an arbitrary rule, it's implied by the form of the
quantum Hamiltonian! Rejecting it destroys unitarity
and then you get faster-than-light signaling!
And cats are \emph{COMPLICATED!} A human mind can't just visualize
a whole cat's anatomy and, and all the cat biochemistry, and what about the
\emph{neurology?} How can you go on \emph{thinking} using a cat-sized brain?"

Professor McGonagall's lips were twitching harder now. "Magic."

"Magic \emph{isn't enough} to do that! You'd have to be a god!"

Professor McGonagall blinked. "That's the first time I've ever been called
\emph{that}."

A blur was coming over Harry's vision, as his brain started to comprehend what
had just broken. The whole idea of a unified universe with mathematically
regular laws, that was what had been flushed down the toilet; the whole notion
of \emph{physics}. Three thousand years of resolving big complicated things
into smaller pieces, discovering that the music of the planets was the same
tune as a falling apple, finding that the true laws were perfectly universal
and had no exceptions anywhere and took the form of simple mathematics governing the
smallest parts, \emph{not to mention} that the mind was the brain and the brain
was made of neurons, a brain was what a person \emph{was}—

And then a woman turned into a cat, so much for all that.

A hundred questions fought for priority over Harry's lips, and the winner poured
out: "And, and what kind of incantation is \emph{Wingardium Leviosa}? Who
invents the words to these spells, nursery schoolers?"

"That will do, Mr.~Potter," Professor McGonagall said crisply, though her eyes
shone with suppressed amusement. "If you wish to learn about magic, I suggest
that we finalize the paperwork so that you can go to Hogwarts."

"Right," Harry said, somewhat dazed. Paperwork. Some things
never changed, it seemed, even in a world of magic.
He pulled his thoughts together. The March
of Reason would just have to start over, that was all.
They still had the experimental method, and that was the important thing.

"Are you alright, darling?" Mum said, putting a hand on her
husband's shoulder.

Professor Verres-Evans did look rather pale. He patted her
hand. "I{\el} I think so, dear, thank you." He then brought
her hand to his lips in a rare show of public affection.
"And{\el} I'm sorry."

Petunia smiled and squeezed his hand. "That's alright. I
was just as doubtful with Lily, and I didn't have half as
many good reasons to be as you."

Dad smiled at her, then looked at Harry. "I apologize to
you too, son. You were right. `The final arbiter is
observation,' indeed. I don't know if I can quite take all
this in properly, but{\el}"

Harry had choked up a bit, and now he smiled back at his
parents. "I had some help, or I would probably have been
just as doubtful. Maybe it's a wizard thing. I'll explain
another time." He turned to Professor McGonagall, who he
now remembered was also the Deputy Headmistress to
Hogwarts. A real school for wizards and witches. He
couldn't begin to imagine what it would be like, with
professors like this. "I'm ready. How do I get to Hogwarts?"

A choked laugh escaped Professor McGonagall, as if extracted from her by
tweezers. "I won't be whisking you away by
magic, if that's what you're expecting. As the letter said,
term starts September 1. I will come again and explain
how transportation will occur, as well as help you obtain
your school supplies."

"Hold on a moment, Harry," his father said. "Remember why you haven't been
going to school up until now? What about your condition?"

Professor McGonagall spun to face Michael. "His condition? What's this?"

"I don't sleep right," Harry said. He waved his hands helplessly. "My sleep
cycle is twenty-six hours long, I always go to sleep two hours later, every
day. I can't fall asleep any earlier than that, and then the next day I go to
sleep two hours later than \emph{that}. 10\PM, 12\AM, 2\AM, 4\AM, until it goes
around the clock. Even if I try to wake up early, it makes no difference and
I'm a wreck that whole day. That's why I haven't been going to a normal school
up until now."

"One of the reasons," said his mother. Harry winced. He
didn't want his potential future teacher and Deputy
Headmistress to have a biased opinion of him.

\emph{Even if it might be a bit deserved?} asked his Inner Critic.

\emph{It could be important for the teachers to know,}
commented his utilitarian side. \emph{Remember when our science project—}

\emph{Shut up or they might not teach us magic!} said his Id,
and the other parts of Harry promptly fell into agreed silence.

McGonagall gave a long \emph{hmmmmm.} "I can't recall hearing about such a
condition before{\el}" she said slowly. "I'll check with Madam Pomfrey to
see if she knows any remedies." Then her face brightened. "No, I'm sure this
won't be a problem—I'll find a solution in time. Now," and her gaze sharpened
again, "what are these \emph{other} reasons?"

Harry sent his parents a glare, then straightened his
shoulders. "I," he said with deliberate gravity, "am a
conscientious objector to child conscription. On grounds that
I should not have to suffer for a disintegrating school
system's failure to provide teachers or study materials of
even minimally adequate quality."

Both of Harry's parents burst out laughing.
"Oh," said Harry's father, eyes bright, "is \emph{that} why you
bit a math teacher in third year?"

"\emph{She didn't know what a logarithm was!}"

"Of course," seconded Mum. "And biting her was a very mature response."

Dad nodded. "A well-considered policy to address the
failings of a disintegrating school system."

"I was \emph{seven years old!} How long are you going to keep on bringing that
up?"

"I know," said his mother sympathetically, "you bite \emph{one} teacher
and they never let you forget it, do they?"

Harry turned to Professor McGonagall as his father
chuckled. "Are you sure you can't just whisk me away now?"

"Quite sure." Professor McGonagall's restrained smile
threatened to burst into a grin at any moment. "And there
is to be no biting of teachers at Hogwarts, is that quite
clear, Mr. Potter?"

Harry scowled at her. "Fine, I won't bite anyone who
doesn't bite me first."

"Better not ask him to build a volcano either," Dad
suggested, as his mother began howling with laughter.
"Not unless this school of yours is magically fireproof."

"\emph{Dad!}" Harry yelled, cheeks burning.

"Well{\el}" Professor McGonagall said. "I think, under the
circumstances, that I should avoid taking you to purchase
your study materials until a day or two before school begins."

"What? \emph{Why?} The other children already know magic, don't
they? I have to start catching up right away! I promise
not to burn down the school!" It occurred to him exactly a
second after saying it out loud that his having to say it
was not particularly encouraging.

"Rest assured, Mr.~Potter," replied Professor McGonagall,
"Everyone at Hogwarts will begin with the basics, and the
school is quite capable of teaching its students without risk
of self-destruction. On the other hand, I suspect that if I
leave you alone for two months with your schoolbooks,
even without a wand, I will return to this house only to
find a crater billowing purple smoke, a depopulated city
surrounding it, and a plague of flaming zebras terrorizing
what remains of Oxford."

Harry's mother and father nodded in perfect unison.

"\emph{Mum! Dad!}"
