% Updated to an edited version of Daystar's Remix of Rationality
% https://www.fanfiction.net/s/9676374/1/Daystar-s-Remix-of-Rationality

\chapter{A Day of Very Low Probability}

\lettrine{H}{arry} James Potter-Evans-Verres was doing his best to
ignore the yelling outside his cupboard.

It was an hour before supper, and he was lying in the
cupboard under the stairs, reading a fantasy novel.
Normally, Harry enjoyed reading in companionable silence with
his father in his study, or tuning out the sound of his
mother's soap operas in the living room, but when he
wanted quiet that even his room couldn't provide, he
would go under the stairs. It was a private, cozy place,
mostly insulated from the sounds of phone conversations,
television, or outside traffic.

This particular night, however, the walls were no match for
the steadily rising voices of Michael Verres-Evans and Petunia
Evans-Verres, and soon Harry began to catch bits and
pieces of the conversation.

\emph{"{\el} just rubbish{\el} fourth time this week{\el} a silly prank,
Petunia—"}

Harry adjusted his glasses and tried to concentrate on the
book. The author was attempting to explain, through an
old wizard's limited grasp of biology and chemistry, how
the dragons in his world breathed fire. Though Harry
generally preferred science fiction, he always enjoyed
fantasy best when the writers at least \emph{tried} to put some
of the magic in rational, understandable terms: it fired up
his imagination to think outside the box for what was
possible, if not terribly probable.

\emph{"{\el} not a prank, I told you{\el} have to show him, or they'll
keep{\el} more and more of them{\el}"}

\emph{"{\el} nonsense, there's no need{\el} worry about crackpots
sending him letters!"}

Unfortunately, now his imagination was preoccupied with
what kinds of letters his dad was keeping from him. Harry
closed his book, no longer able to concentrate as a familiar
bitterness flared up in him.

It wasn't that his parents mistreated him. Far from it—he'd
been sent to the best primary schools, and when that
proved insufficient he was given the best tutors an endless
pool of starving university students could provide. He'd
always been encouraged to study whatever caught his
attention, was bought all the books he wanted, was
sponsored in whatever math or science competitions he
entered. Harry knew he was exceedingly lucky, and he was
always grateful for what his parents gave him{\el} but he
would have been satisfied with half as much if it meant he
had their respect.

Of course, if asked, his parents would say they respected
him. An Oxford Professor of Biochemistry and his liberal
wife were \emph{expected} to show an enlightened view of
child-rearing that included respect. But that respect meant
something different than it would for a fellow adult, who
they would never have dreamed of talking about as if he
weren't in the house, let alone making decisions for him.

It wasn't their fault: society as a whole had such low
expectations of children. And if it was ever going to
change, it would be up to those like him to change it.

So Harry swung his legs out of the small hammock he'd
strung to the walls, turned off the lantern his father had
hung up for him, and opened the door into the hallway.

The voices immediately quieted. By the time he stepped
into the living room, his parents were sitting calmly on the
couch, watching the news on a television that stuck out
from its surroundings. Their living room was
dominated by books. Every inch of wall space was covered
by a bookcase going almost to the ceiling. Some
bookshelves were stacked to the brim with hardback
books: science, math, history, and everything else. Other
shelves had two layers of paperback science fiction, one
set right side up, the other stacked sideways in what was left
of the space above. And it still wasn't enough. Books 
overflowed onto the tables and the sofas, covered the top
of the television, and made little stacks under the windows.

"Hi, Mum, Dad. Is everything alright?"

"Hello, Harry." His mother turned to him with a warm
smile, her face still young and pretty despite her age. "Yes,
everything's fine."

"Did we disturb your reading, son?" his father said, looking
contrite. "We're sorry, our debate got a bit passionate at
the end there," Michael chuckled.

Harry and his mother exchanged knowing smiles. Professor
Verres-Evans viewed arguments as uncivilized, and so any
he participated in were automatically elevated in status to
"debate."

"It's alright. I just couldn't help overhearing,"
Harry said with mild emphasis, "and it sounded like a letter
arrived for me?"

He saw it in the quick glance they gave each other—his
mother's expectant, his father's calculating. Harry knew his
father was struggling with some mighty cognitive
dissonance. One part of him felt guilty from withholding
someone's mail from them, a grievous breach of privacy.
The other part felt entitled by societal norms that parents
were allowed to decide for their children what information
they should or shouldn't have, no matter how bright and
precocious those children might be.

"Yes," Petunia said after the silence stretched on a few
seconds. "It's the first time I've seen it, or I would have
told you sooner. Your father thinks it's just prank mail, but
he doesn't understand—"

"Well, no harm in having a look then, right?" Harry said.
He held his hand out expectantly, brow raised in an
expression of innocent patience. He wasn't quite sure what
he'd do if his father refused—trying to reason with him
rarely worked on any topic that concerned Harry's
subordinate status.

After a moment, though, his father nodded and stood up,
walking toward the trash and fishing an envelope and a
couple papers from it. "Quite right, Harry, no harm in
looking. You're a bright boy, and I know you won't get
suckered in by whatever crock they're selling."

Michael handed the letters and envelope to Harry, who
had to choke back a retort to the patronizing tone his
father had adopted now that he was giving in. Admitting
one's mistakes was for scientific journals, apparently{\el} not
for adults to do to children.

Harry chided himself on such bitter thoughts as he went
to the table. He knew this was a sore spot for him, and it
occasionally took a while for his temper to calm down. So
he forced himself to smile back at his dad, then
straightened out the first thick paper and began to
read, acutely aware of his parents' stares.
Harry's eyes scanned the letter in a few seconds, blinked,
then looked up to meet theirs.

"What."

Michael Evans-Verres smiled. "Yes, rather silly I th—"

Harry held up his hand, then looked back down at the
parchment (simple "paper" didn't suffice to describe this material)
and slowly reread the message.

\begin{writtenNote}
\letterAddress{Dear Mr.~Potter,}

We are pleased to inform you that you have been
accepted at Hogwarts School of Witchcraft and Wizardry.
Please find enclosed a list of all necessary books and
equipment.

Term begins on September 1st. We await your owl by no
later than July 31st.

\letterClosing[Yours sincerely,]{Minerva McGonagall, Deputy Headmistress}
\end{writtenNote}

On the second sheet he found a list that wouldn't be out
of place in a rulebook from a fantasy role-playing game.

"What is it, some kind of late summer camp?" Harry
asked as he eyed the impressive seal heading the
parchment: a lion, snake, raven and badger surrounding
an ornate \emph{H}. He smiled as he looked back at the name of
the school. \emph{Heh. "Hogwarts." What, was "Newteyes" taken?}

"No, Harry," his mother said. "It's not a summer camp. As
I was telling your father{\el}" She took a deep breath,
straightened in her seat, and avoided looking at her
husband, her gaze steady on Harry. "My sister—your mother,
Lily—was a witch. She got that same letter. I'd promised
to keep it secret, my whole family did, but now it's clear
you're meant to know, if they've come for you like they
did for her."

Harry exchanged a glance with his father, feeling a mix of
exasperation and confusion. Mum rarely spoke of his
biological parents. It wasn't taboo or anything, it just never
really came up. They'd died in a car crash when he was
one year old{\el} the same crash which had given him the
lightning shaped scar on his forehead. To hear that they
were Wiccan wasn't terribly surprising considering some of
Petunia's beliefs, but the gravity of her tone didn't match
the subject matter.

"Well that's, er, very interesting, I guess. But what does
her religion have to do with me? Who's `they'?" He didn't
particularly like the ominous sound of them "coming for
him," whoever they were. He imagined a shadowy coven
meeting in a forest and pronouncing it time to bring the
young Potter into the fold.

"It wasn't a religion. I'm saying she was an \emph{actual} witch.
She could do magic. Her husband—your father—he was a
wizard. They both went to this magic school, Hogwarts,
when they were eleven. And since you've received that letter, it
means you're a wizard too, Harry."

Michael Verres-Evans laughed, and Harry almost joined
him. Petunia Evans-Verres had always been something of
the odd-woman-out in their family. Some of the most
"spirited debates" he could remember between his parents
involved her superstitions, and he had a clear childhood
memory of her waving a crystal of some kind in careful
patterns over him when he was sick.

When he had been younger, he used to enjoy going with her to
the smoky, mysterious shops she would occasionally
frequent, with their pungent odors and exotic wares.
Thankfully his father's books had taught him how to
critically examine the beliefs sold in such places, and a few
years ago he had begun to find their air of obscure
mysticism groundless and mildly irritating.

Harry smiled down at the parchment listing the "school
supplies." Wand, spell books, potion ingredients{\el} he quickly
scanned the latter. Nope, no "hog warts" listed, though
newt eyes did indeed show up, as well as powdered hens'
teeth. He wondered how expensive that would be: he
knew there was some research being done on atavism in
chickens that resulted in them growing vestigial teeth, and
that the mutation was rather rare. Aboriginal shaman and
medicine men must have found plenty of uses for it, or
imagined them at any rate. He wondered what Hogwarts
pretended to use them for. Good dental hygiene?

And yet he didn't laugh with his father. Because{\el}

Because somewhere in him was a strange certainty that
she was right, in this, the most unlikely of cases.
\emph{You're a wizard too, Harry.}

"Well, maybe someday he'll be a wizard at chess," his
father said, still smiling as he turned back to the news.
"But if whoever keeps sending those letters shows up at
the door in a robe and pointy hat, I'm calling the men in
the white coats."

Petunia continued to look only at Harry, her gaze intent,
waiting.

"Mum," he said. "What do you mean by `wizard'?"

Petunia bit her lip. "I can't just tell you. You'll think I'm—"
She swallowed, and Harry felt confused. His mother had
always defended her less rational beliefs with
exasperating calmness, merely shrugging off logical arguments
and relying on some inner conviction. This sudden
nervousness, and the confusion he felt from it, made him
pay attention. "Listen. I wasn't{\el} always like this{\el}" She
gestured at herself, as though to indicate her lithe form.

"Lily did this. Because I{\el} I \emph{begged} her. For years, I
begged her. Lily had always been prettier than me, and
I'd{\el} been mean to her, because of that, and then she
got \emph{magic}, can you imagine how I felt? And I \emph{begged} her
to use some of that magic on me so that I could be
pretty too, even if I couldn't have her magic, at least I
could be pretty."

Harry watched in alarm as tears gathered in Petunia's eyes.

"And Lily would tell me no, and make up the most
ridiculous excuses, like the world would end if she were
nice to her sister, or a centaur told her not to—the most
ridiculous things, and I hated her for it. And when I had
just graduated from university, I was going out with this
boy, Vernon Dursley{\el} he was fat, and he was the only boy
who would talk to me. And he said he wanted children,
and that his first son would be named Dudley. And I
thought to myself, \emph{what kind of parent names their child
Dudley Dursley?} It was like I saw my whole future life
stretching out in front of me, and I couldn't stand it. And
I wrote to my sister and told her that if she didn't help
me I'd rather just—"

Petunia stopped. Harry felt somewhat wretched for being
responsible for her having to relate such an obviously
painful memory. A glance at his father showed his dad
similarly stricken. He'd never known that Mum had been
through such a dark period, had been so envious of her
sister{\el} he wondered how much guilt she must have felt
after his biological parents had died.

"Anyway," Petunia said, her voice small, "Lily gave in. She
warned me it was dangerous, and I said I didn't care. I
drank this potion and I was sick for weeks, but when I
got better my skin cleared up and I finally filled out and{\el}
I was beautiful. People were \emph{nice} to me," her voice broke,
"and after that I couldn't hate my sister any more,
especially when I learned what her magic brought her in
the end—"

"Darling," Michael said gently, "you got sick, you gained
some weight while resting in bed, and your skin cleared up
on its own. Or being sick made you change your diet—"

"No, it was nothing like that," Petunia said. "It was magic,
real magic. I saw it, other things—"

"Petunia," Michael said. The annoyance was creeping back
into his voice. "You \emph{know} that can't be true. Do I really
have to explain why?"

Petunia wrung her hands. She seemed to be on the verge
of tears. "My love, I know I can't win arguments with you,
but please, you have to trust me on this—"

\emph{"Dad! Mum!"}

The two of them stopped and looked at Harry. He took a
deep breath and thought about the problem. "Mum, your
parents didn't have magic, did they?"

"No," Petunia said. "Just Lily."

"Then your family also must not have believed her letter.
How did \emph{they} get convinced?"

"Ah{\el}" Petunia said. "They didn't just send a letter. They
sent a professor from Hogwarts. He—" Petunia's eyes
flicked to Michael. "He showed us some magic."

"Well, there we are then. You don't have to fight over
this," Harry said firmly. "If it's true, we can just get a
Hogwarts professor here and see the magic for ourselves,
and Dad will admit that it's true. And if not, then Mum will
admit that it's false. That's what the experimental method
is for, so that we don't have to resolve things just by
arguing." Hoping against hope that this time, just this once,
they might listen to him{\el}

"Oh, come now, Harry," Professor Verres-Evans said.
"Really, \emph{magic}? I thought you'd know better than to take
this seriously, even if you're only ten."

\emph{I. Shall. SCREAM.}

"Mum," Harry said instead, keeping his voice calm. "If you
want to win this argument with Dad, look in chapter two
of the first book of the \emph{Feynman Lectures on Physics}.
There's a quote there about how philosophers say a great
deal about what science absolutely requires, and it's all
wrong, because the only rule in science is that the final
arbiter is observation—that you just have to look at the
world and report what you see. Um{\el} off the top of my
head I can't think of where to find something about how
it's an ideal of science to settle things by experiment
instead of arguments—"

His mother looked at him and smiled. "Thank you, Harry.
But," she looked back at her husband. "I don't want to
win an argument with your father. I want my husband to
just{\el} listen to his wife who loves him, and trust her just
this once{\el}"

Harry closed his eyes briefly. \emph{Hopeless}. Both his parents
were hopeless.

Now they were getting into one of \emph{those} arguments again,
one where his mother tried to make her husband feel
guilty, and his father tried to make his wife feel stupid.

"I'm going to go to my room," Harry announced. His voice
trembled a little. "Please try not to fight too much about
this, Mum, Dad, we'll know soon enough how it comes out, right?"

"Of course, Harry," said his father, and his mother gave
him a reassuring kiss, and then they went on "debating"
while Harry climbed the stairs to his bedroom.

He shut the door behind him and tried to think, wandering
past his own bookshelves crammed with textbooks and
sci-fi to lie on his bed.

The funny thing was, he \emph{should} have agreed with Dad. No
one had ever seen any evidence of magic, and according
to Mum, there was a whole magical world out there. How
could anyone keep something like that a secret in a world
of video cameras and spy satellites? More magic? That
seemed like a rather suspicious sort of excuse.

Except that some part of Harry was utterly convinced that
what his Mum said was true. He was magic{\el} a wizard.

Was it simple ego? What child didn't want to believe they
possessed hidden, magic powers? He knew he had an
inflated sense of self-importance as others judged it. He'd
always vowed to one day justify it by proving himself
unique. Of course, he'd figured it would be somewhere in
the realm of science. He'd imagined becoming a world
renowned biologist, curing cancer and extending lifespans
indefinitely. Or going into physics to perfect cold fusion,
ending the planet's energy needs and propelling humanity
to the stars. Reasonable things. Mostly. Not magic, at any
rate.

Maybe his powers of reason had been impaired somehow.
He frowned, probing his skull with his fingers as if some
wound would present itself. He hadn't hit his head on
anything lately{\el} not that he could remember in any case.
\emph{Would} he even remember? There was a scary thought.
Harry mentally jumped through some quick mental hoops
to confirm that yes, the least complicated answer that fits all
the facts is most likely to be the true one, that all claims
require evidence and that extraordinary claims require
extraordinary evidence, that two plus two is still four.

It should have been a clean case for Mum joking, lying or
being insane, in ascending order of awfulness. If Mum had
sent the letter herself, that would explain how it arrived at
the letterbox without a stamp. A little insanity was far, far
less improbable than the universe really working like the
contents of that letter implied.

What about his mother's other views? Was he any more
susceptible to those? He considered her belief that atoms
arranged in a particular pattern identified as a "crystal"
could somehow destroy bacteria or viruses in his body
when touched to his skin{\el} specifically those bacteria or
viruses deemed "harmful," opposed to all the beneficial
ones{\el} Yes, that he could still rationally reject as a form of
wish fulfillment without any evidence to back it up. If the
person from Hogwarts came to their house and started
bending spoons, he would toss the letter in the trash and
think nothing further of it.

But that \emph{he} was magical{\el} that irrational belief still stayed.
And he could think of no evidence to account for it{\el} no
moments in his life when he'd exhibited supernatural or
inexplicable powers, no hidden talent manifesting in times
of great peril or passion. Yet he still believed he had magic.

Harry rubbed his forehead, grimacing. \emph{Don't believe
everything you think}, Harry reminded himself. \emph{So where
do you come from, strange little prediction? Why do I
believe what I believe?}

Usually Harry was pretty good at answering that question,
but in this particular case, he had no \emph{clue} what his brain
was thinking. He hadn't had a belief so clearly based on
faith since he was very young. Some people, unfamiliar
with rationalism or the scientific method, seemed to think
that science took faith, since no one did every experiment
themselves, but rather relied on other scientists or
textbooks to tell them what was true or not true.

The problem with this view was that no scientist had
"faith" in textbooks, other scientists, or even the scientific
method. They had \emph{confidence} in them. Somewhere,
someone was able to do the experiments, verified the
results through repeated tests, and then subjected their
findings to peer review so others could repeat the
experiments. And if he wanted, Harry could take the time
and effort to learn the information and repeat the
experiment himself. Belief in science relied on the \emph{external},
not the internal, and thus could be shown to others,
taught and learned. He no more had faith in science than
he had faith that Dad's car would start tomorrow: he had
confidence based on experimentation and observation.

This new belief, however, was not based on external
factors. He couldn't describe it to anyone in a way that
would make sense. He couldn't demonstrate the belief and
have it peer reviewed. It just was.

Harry mentally shrugged. A button calls to be pushed, a
handle yearns to be turned, and the thing to do with a
testable hypothesis is to go and test it.

He went to his desk, shoved some of the books to the
side, took a piece of lined paper from a drawer, and began writing.

\begin{writtenNote}
\letterAddress{Dear Deputy Headmistress}
\end{writtenNote}

Harry paused, reflecting; then he discarded the paper for another, tapping another
millimeter of graphite from his mechanical pencil. This called for careful
calligraphy.

\begin{writtenNote}
\letterAddress{Dear Deputy Headmistress Minerva McGonagall,}

\letterAddress{Or Whomsoever It May Concern:}

I recently received your letter of acceptance to Hogwarts, addressed to
Mr.~H.~Potter. You may not be aware that my genetic parents, James Potter and
Lily Potter (formerly Lily Evans) are dead. I was adopted by Lily's sister,
Petunia Evans-Verres, and her husband, Michael Verres-Evans.

I am extremely interested in attending Hogwarts, conditional on such a
place actually existing. Only my mother Petunia says she knows about magic, and
she can't use it herself. My father is highly skeptical. I myself am uncertain.
I also don't know where to obtain any of the books or equipment listed in your
acceptance letter.

Mother mentioned that you sent a Hogwarts representative to Lily Potter
(then Lily Evans) in order to demonstrate to her family that magic was real,
and, I presume, help Lily obtain her school materials. If you could do this for
my own family it would be extremely helpful.

\letterClosing[Sincerely,]{Harry James Potter-Evans-Verres.}
\end{writtenNote}

Harry added their current address, then folded up the letter and put it in an
envelope, which he addressed to Hogwarts. Further consideration led him to
obtain a candle and drip wax onto the flap of the envelope, into which, using a
penknife's tip, he impressed the initials H.J.P.E.V. If he was going to descend
into this madness, he was going to do it with style.

Then he opened his door and went back downstairs. His
father was sitting in the living-room and reading a book of
higher mathematics to show how smart he was, and his mother
was in the kitchen preparing one of his father's favorite
meals to show how loving she was. It didn't look like they
were talking to one another at all. As scary as arguments
could be, \emph{not arguing} was somehow much worse.

"Mum," Harry said into the unnerving silence, "I'm going to
test the hypothesis. According to your theory, how do I
send a letter to Hogwarts?"

His mother turned from the sink to look at him uncertainly.
"I don't know{\el} I think you have to own a magic owl."

That should have sounded highly suspicious, \emph{oh, so there's
no way to test your theory then}, but the peculiar
certainty in Harry seemed willing to stick its neck out even further.

"Well, the letter got here somehow," Harry said, "so I'll
just wave it around outside and call `letter for Hogwarts!'
and see if an owl picks it up. Dad, do you want to come and watch?"

His father shook his head minutely and kept on reading.
\emph{Of course}, Harry thought to himself. Magic was a
disgraceful thing that only stupid people believed in; if his
father went so far as to \emph{test} the hypothesis, or even
\emph{watch} it being tested, that would feel like \emph{associating}
himself with that{\el}

Only as Harry stumped out the back door into the garden
did it occur to him that if an owl \emph{did} come down and
snatch the letter, he was going to have some trouble
telling Dad about it.

\emph{But—well—that can't \emph{really} happen, can it? No matter
what my brain seems to believe. If an owl really comes
down and grabs this envelope, I'm going to have worries
a lot more important than what Dad thinks.}

Harry took a deep breath and raised the envelope into the air.

He swallowed.

Calling out "\emph{Letter for Hogwarts!}" while holding an envelope
high in the air in the middle of your own back garden
was{\el} actually pretty embarrassing, now that he thought
about it.

\emph{No. I'm better than this. I will use the scientific method
even if the result makes me feel stupid.}

"Letter—" Harry said, but it actually came out as more of a
whispered croak.

Harry steeled his will and shouted into the empty sky,
"\emph{Letter for Hogwarts! Can I get an owl?}"

"Harry?" asked a bemused woman's voice from nearby.

Harry yanked his hand down as if it had caught fire,
hiding the envelope behind his back. His whole face was hot with shame.

An old woman's face peered out from over the
neighboring fence, her grizzled grey hair escaping from her
hairnet. Mrs.~Figg, the occasional babysitter. "What are you
doing, Harry?"

"Nothing," Harry said in a strangled voice. "Hi, Mrs.~Figg.
I'm just{\el} testing a really silly theory—"

"Did you get your acceptance letter from Hogwarts?"

Harry froze.

"Yes," Harry's lips said a little while later. "I got a letter
from Hogwarts. They say they want my owl by July 31st, but—"

"But you don't \emph{have} an owl. Poor dear! I can't imagine
\emph{what} someone must have been thinking, sending you just
the standard letter."

A wrinkled arm stretched out over the fence and opened
an expectant hand. Hardly thinking himself at this point,
Harry gave over his envelope.

"Just leave it to me, dear," said Mrs.~Figg, "and in a jiffy
or two I'll have someone over."

And her face disappeared from over the fence.

There was a long silence in the garden.

Then Harry's voice said, calmly and quietly, "What."
