\chapter{Comparing Reality To Its Alternatives}

\lettrine{P}{etunia} Evans-Verres looked at Harry in the rear-view
mirror of the car as if she could easily guess his thoughts,
and seemed troubled by them. He'd spent the past few
weeks in a mild frenzy, first interrogating her for what little
direct experience she had about magic ("No, Mum, tell me
what you've \emph{seen}, not what you've guessed or read
about.") then doing independent research, which had
quickly proven fruitless. Any books on magic he found
involved complicated rituals to bring about some minor,
vague misfortune, or wishing yourself riches and happiness
through "positive attraction," or some other such
unfalsifiable feel-good fluff. Nothing \emph{remotely} close to lifting
a man off the ground with a couple words and the wave
of a stick, let alone turning into a cat, and no mention of
a "Hogwarts" anywhere.

Clearly all the \emph{real} magic was kept out of bookstores or
libraries by some organized effort, a notion he found both
troubling and thrilling. On the one hand, he was about to
be part of a massive, worldwide conspiracy the likes of
which he'd only read about in fiction. On the other hand,
the reality of a group of people capable of secretly
enforcing such a conspiracy was mildly terrifying. He
wondered how omnipotent they really were, and whether
non-magic authorities were involved in the cover-up.

The day before yesterday, another message had
arrived at their house. Professor McGonagall had supplied
them with a time and an address where Harry could meet
her to obtain his school supplies. So this morning, Harry's
mother had driven him to London, uncharacteristically quiet
and nervous. Harry assumed she was worried he would
make a bad impression, but he was determined not to get
into any trouble that might jeopardize his acceptance into
the magical world. If the past few weeks had confirmed
anything about his nature to himself, it was that he
couldn't stand being aware of a mystery and not having
the means to solve it. Just imagining going on with his life
without learning more about magic{\el} any scientific field he
went into would drive him mad as he considered the true
nature of reality that he'd caught a glimpse of.

Once they arrived at the appropriate address, Harry's
mother parked beside a row of shops. Harry stepped out
of the car and looked around, and his mother rolled down
her window.

"Well," Petunia said after a pause, looking up and down
the sidewalk. "I don't see Professor McGonagall{\el} though
we are a bit early. Where do you suppose the place is?
`The Leaky Cauldron,' wasn't it?"

Harry turned in a slow circle, scanning the shops along the
street. Nothing looked like a place that would sell magic
wands, even as a joke. There was a fashionable clothing
store, a hair boutique, an ice cream parlor, some fast food
restaurants, a book shop (which he quickly jogged into,
looked around a bit, then left), a pub—"There," he said,
pointing to The Leaky Cauldron, a quaint brick building
tucked between the book shop and a record store. "Maybe
she's already inside."

"Hm?" Harry's mother looked vaguely in the direction he'd
pointed. "Did you say you saw her?"

Harry began to point again, then stopped and looked at
his Mum, then back at the pub. She was looking right at
it. "What do you see between that bookshop and the
record store?"

"What do you mean, dear? In the alley?"

\emph{Alley?} From Harry's perspective, the walls of The Leaky
Cauldron were pressed up against its neighbors. "You don't
see the pub right there?" he asked, pointing straight at it
again.

"No," Petunia said. "You mean to tell me there is one?"

Harry felt an electric thrill go up his spine and simply
couldn't help himself. He approached a nearby couple.
"Excuse me. I'm afraid I brought
the wrong prescription with me this morning and can't
quite make out the store signs. Could you read them to
me please, from right to left?" He gestured.

The man gave him a curious look, but the woman began
listing names. Harry watched her eyes as she named the
book shop, then the record store, without mentioning the
Leaky Cauldron. It looked as though her gaze simply
passed over where it was without registering it.

"Thank you." He returned to his mother, shifting his weight
from foot to foot with nervous energy as he stood beside
the car and examined the pub. "It's not just you. They
couldn't see it either." Here it was. Proof, however subtle,
that he wasn't like other people. The now-familiar sense of
disorientation came over him, and his mind raced with
possibilities for how the cloaking worked. Harry wondered
what would happen if he threw a rock out of the pub's
window. Would the glass suddenly become visible to the
people on the street? It was all he could do to not rush
to the pub and begin experimenting with his mother's
perception.

If wizard folk could do things like \emph{this}, it was no wonder
Harry couldn't find any books about them. He wondered
now if a massive conspiracy was really needed to hide
records of the magical world. What if non-magic folk
couldn't even see the books? What else had he seen in
his life without realizing no one else could? Maybe some
other safety mechanism was in place, like he needed to
know the name of the pub as well—

"Good morning, Mr.~Potter."

Harry spun around to see Professor McGonagall, who
seemed utterly unconcerned with the odd
looks she was getting from passersby.

"Good morning, Professor," Harry said. "Why can't Mum
see The Leaky Cauldron?"

"It's enchanted against Muggle notice." Professor McGonagall
turned to his mother. "Good morning, Mrs.~Evans-Verres.
I'm sorry to have kept you waiting."

"Not at all, we'd just arrived." Petunia looked back at
Harry, with the same nervous air she'd had all day.

"Well, I'll be back this evening to pick you up. Be good, Harry."

She kissed him goodbye and drove away. Harry watched his mother
go, then turned to Professor McGonagall. "What's a `Muggle?'"

Professor McGonagall's lips twitched. "It's good to see you
again too, Mr.~Potter. Muggles are what we call those
without a drop of magic in them. Shall we?"

Harry followed her toward the pub. "Okay, so Dad's a
Muggle, but Mum too? Her sister was a witch, doesn't that
mean she had some magic in her family?"

"Oh, no, if her parents were both Muggles then she's
a Muggle as well," Professor McGonagall explained in her prim,
Scottish voice. "What you're thinking of is what we call a
Squib. Though children of a witch and wizard, they cannot
perform magic themselves, but they can perceive
many magical things and use enchanted items."

Harry was trying to filter this information through his
understanding of genetics, but he was distracted partway as
they entered The Leaky Cauldron. He whipped his head
around to see if anything of note happened as they did,
but he couldn't detect any invisibility field descending on him,
and no one in the street seemed to notice two people
suddenly disappear.

The inside of the pub was somewhat dark and shabby, with
wooden tables scattered about the shadows and a grubby
bar that dominated the far wall. About a dozen people
were inside, most dressed in variously colored robes.

"Good morning, Professor McGonagall," said the barman
with a smile.

"Good morning, Tom."

"Is there anything I could get for—Good Lord." The barman
peered at Harry, gaze drawn to his forehead. "is this—can this be—?"

Harry leaned towards the bar of the Leaky Cauldron as best he could, though it
came up to somewhere around the tips of his eyebrows. A question like
\emph{that} deserved his very best.

"Am I—could I be—maybe—you never know—if I'm \emph{not}—but then the
question is—\emph{who?}"

"Bless my soul," whispered the old barman. "Harry Potter{\el} what an
honor."

Harry blinked, then rallied. "Well, yes, you're quite perceptive; most people
don't realize that so quickly—"

"That's enough," Professor McGonagall said. Her hand tightened on Harry's
shoulder. "Don't pester the boy, Tom, he's new to all this."

"But it is him?" quavered an old woman. "It's Harry Potter?" With a scraping
sound, she got up from her chair.

"Doris—" McGonagall said warningly. The glare she shot around the room
was enough to stop most others from doing
more than muttering among themselves and staring;
some paused halfway out of their seats.

"I only want to shake his hand," the woman whispered. She bent low and stuck
out a wrinkled palm, which Harry, feeling confused and more uncomfortable than
he ever had in his life, carefully shook. Tears fell from the woman's eyes onto
their clasped hands. "My grandson was an Auror," she whispered to him. "Died in
seventy-nine. Thank you, Harry Potter. Thank heavens for you."

"You're welcome," Harry said automatically, and then he turned his head and
shot Professor McGonagall a frightened, pleading look.

Others began to approach them again, and Professor
McGonagall slammed her foot down. It made a noise that
gave Harry a new referent for the phrase "Crack of
Doom", and the other bar patrons once again froze in
place just as the general rush was about to start.

"We're in a hurry," Professor McGonagall stated calmly.

They left the bar without any trouble and entered the courtyard
behind it, which was surrounded on all sides by high brick walls.

"Professor?" Harry said, meaning to ask what was going on, but oddly
finding himself asking an entirely different question instead.
"Who was that pale man, by the corner? The man with
the twitching eye, slumped in his seat?"

"Hm?" said Professor McGonagall, sounding a bit surprised; perhaps she hadn't
expected that question either. "That was Professor Quirinus Quirrell. He'll be
teaching Defense Against the Dark Arts this year at Hogwarts."

"I had the strangest feeling that I knew him{\el}" Harry rubbed his
forehead. "And that I shouldn't ought to shake his hand." Like meeting someone
who had been a friend, once, before something went drastically wrong{\el}
that wasn't really it at all, but Harry couldn't find words. "And what
\emph{was}{\el} all of that?"

Professor McGonagall was giving him an odd glance. "Mr.~Potter{\el} do you
know{\el} how \emph{much} have you been told{\el} about how your parents
died?"

Harry returned a steady look. "My parents are alive and
well, thank you. They've told me that my \emph{genetic} parents
were killed in a car accident when I was one year old."

"An admirable loyalty," said Professor McGonagall. Her voice went low. "Though
it hurts a little to hear you say it like that. Lily and James were friends of
mine."

Harry looked away, suddenly ashamed. "I'm sorry," he said in a small voice.
"But I \emph{have} a Mum and Dad. And I know that I'd just make myself unhappy
by comparing that reality to{\el} something perfect that I built up in my
imagination."

"That is amazingly wise of you," Professor McGonagall said quietly. "But your
\emph{genetic} parents died very well indeed, protecting you."

\emph{Protecting me?}

Something strange clutched at Harry's heart. "So it{\el}
wasn't a car crash? What \emph{did} happen?"

Professor McGonagall sighed. Her wand tapped Harry's forehead, and his vision
blurred for a moment. "Something of a disguise," she said, "so that this
doesn't happen again, not until you're ready." Then her wand licked out again,
and tapped three times on a brick in the wall{\el}

{\el} which hollowed into a hole that dilated and expanded and
shivered into a huge archway, revealing a cobbled street full of
odd shops advertising everything from actual \emph{cauldrons} to
\emph{dragon liver.}

Harry didn't blink. It wasn't like anyone was turning into a cat.

"Welcome, Mr.~Potter, to Diagon Alley."

And they walked forwards, together, into the wizarding world.

Here, Harry was sure, was the true testament to the
effectiveness of magical secrecy. A whole long, winding
street of London City completely unknown by its
inhabitants. Only powerful magic or political agreements of
the highest order could keep airplanes or satellites from
taking note of such a place. Here were merchants hawking
Bounce Boots ("Made with real Flubber!"). There were
goggles that would turn anything you looked at green, and
a lineup of comfy armchairs with ejection seats for
emergencies. Some of the buildings were merely a story
or two high, while others had multiple floors and were
oddly structured, as though relying on magic to keep them upright.

Harry's head kept rotating as if trying to wind itself off his
neck. It was like walking through the magical items section of an
\emph{Advanced Dungeons and Dragons} rulebook (he enjoyed
reading those, even though he had no one
to play the games with). Harry desperately didn't want to miss a
single item for sale, in case it was one of the three you needed to complete
the cycle of infinite \emph{wish} spells.

Then Harry spotted something that made him unthinkingly
veer off from the Deputy Headmistress.
He was heading straight into the shop, a front of blue bricks
with bronze-metal trim, when Professor McGonagall's voice
brought him back to reality.

"Mr.~Potter?" she said.

Harry blinked, then realized what he'd just done. "I'm sorry! I forgot for a
moment that I was with you instead of my family." Harry gestured at the shop
window, which displayed fiery letters that shone piercingly bright and yet
remote, spelling out \emph{Bigbam's Brilliant Books}. "When you walk past a
bookshop you haven't visited before, you have to go in and look around. That's
the family rule."

"That is the most Ravenclaw thing I have ever heard."

"What?"

"Nothing. Mr.~Potter, our first step is to visit Gringotts, the bank of the
wizarding world. Your \emph{genetic} family vault is there, with the
inheritance your \emph{genetic} parents left you, and you'll need money for
school supplies." She sighed. "And, I suppose, a certain amount of spending
money for books could be excused as well. Though you might want to hold off for
a time. Hogwarts has quite a large library on magical subjects. And the tower
in which I strongly suspect you will be living has a more broad-ranging
library of its own. Any book you bought now would probably be a duplicate."

Harry nodded, and they walked on.

"Don't get me wrong, it's a \emph{great} distraction," Harry said as his head
kept swiveling, "probably the best distraction anyone has ever tried on me,
but don't think I've forgotten about our pending discussion."

Professor McGonagall was silent for a time.
"Your parents—or your mother at any rate—may
have been very wise not to tell you."

"So you wish that I could continue in blissful ignorance? There is a certain
flaw in that plan, Professor McGonagall."

"I suppose it would be rather pointless," the witch said tightly, "when anyone
on the street could tell you the story. Very well."

And she told him of He-Who-Must-Not-Be-Named, the Dark Lord, Voldemort.

"Voldemort?" Harry whispered. It should have been funny, but it wasn't. The
name burned with a cold feeling, ruthlessness, diamond clarity, a hammer of
pure titanium descending upon an anvil of yielding flesh. A chill swept over
Harry even as he pronounced the word, and he resolved then and there to use
safer terms like You-Know-Who.

The Dark Lord had raged upon wizarding Britain like a wilding wolf, tearing and
rending at the fabric of their everyday lives. Other countries had wrung their
hands but hesitated to intervene, whether out of apathetic selfishness or
simple fear, for whichever was first among them to oppose the Dark Lord, their
peace would be the next target of his terror.

(\emph{The bystander effect,} thought Harry, thinking of Latane and Darley's
experiment, which had shown that you were more likely to get help if you had an
epileptic fit in front of one person than in front of three. \emph{Diffusion of
responsibility, everyone hoping that someone else would go first.})

The Death Eaters had followed in the Dark Lord's wake and in his vanguard,
carrion vultures to pick at wounds, or snakes to bite and weaken. The Death
Eaters were not as terrible as the Dark Lord, but they were terrible, and they
were many. And the Death Eaters wielded more than wands; there was wealth
within those masked ranks, and political power, and secrets held in blackmail,
to paralyze a society trying to protect itself.

An old and respected journalist, Yermy Wibble, called for increased taxes and
conscription. He shouted that it was absurd for the many to cower in fear of
the few. His skin, only his skin, had been found nailed to the newsroom wall
that next morning, next to the skins of his wife and two daughters. Everyone
wished for something more to be done, and no one dared take the lead to propose
it. Whoever stood out the most became the next example.

Until the names of James and Lily Potter rose to the top of that list.

And those two might have died with their wands in their hands and not regretted
their choices, for they \emph{were} heroes; but they had an infant
child, their son, Harry Potter.

Tears were coming into Harry's eyes. He wiped them away in anger or maybe
desperation, \emph{I didn't know those people, not really, they aren't my
parents} now, \emph{it would be pointless to feel so sad for them—}

When Harry was done sobbing into the witch's robes, he looked up, and felt a
little bit better to see tears in Professor McGonagall's eyes as well.

"So what happened?" Harry said, his voice trembling.

"The Dark Lord came to Godric's Hollow," Professor McGonagall said in a
whisper. "You should have been hidden, but you were betrayed. The Dark Lord
killed James, and he killed Lily, and he came in the end to you, to your cot.
He cast the Killing Curse at you, and that was where it ended. The Killing
Curse is formed of pure hate, and strikes directly at the soul, severing it
from the body. It cannot be blocked, and whomever it strikes, they die. But you
survived. You are the only person ever to survive. The Killing Curse rebounded
and struck the Dark Lord, leaving only the burnt hulk of his body and a scar
upon your forehead. That was the end of the terror, and we were free. That,
Harry Potter, is why you are often called `The Boy Who Lived,'
and why people want to see the scar on your forehead, and why they
want to shake your hand."

The storm of weeping that had washed through Harry had used up all his tears;
he would not cry again.

(And somewhere in the back of his mind was a small, small note of confusion, a
sense of something wrong about that story; and it should have been a part of
Harry's art to notice that tiny note, but he was distracted. For it is a sad
rule that whenever you are most in need of your art as a rationalist, that is
when you are most likely to forget it.)

Harry detached himself from Professor McGonagall's side. "I'll—have to think
about this," he said, trying to keep his voice under control. He stared at his
shoes. "Um. You can go ahead and call them my parents, if you want, you don't
have to say `genetic parents' or anything. I guess there's no reason I can't
have two mothers and two fathers."

There was no sound from Professor McGonagall.

And they walked together in silence, making their way through the enchanted street.
