\chapter{The Efficient Market Hypothesis}

\lettrine{G}{ringotts} Bank turned out to be an imposing
building made of snow-white marble. It was located
partway down Diagon Alley, near an intersection with
something called Knockturn Alley, and it towered over the
neighboring shops. Its architecture seemed subtly different
than the "Muggle" British buildings the wizarding world
seemed to mimic, but Harry had never studied architecture
enough to pinpoint how.

He was also too distracted by the pair of goblins standing
at the bank's ornate double doors.

They were dressed in perfectly tailored uniforms of scarlet
and gold, and discreetly examined everyone that walked by
the bank. Harry knew they were goblins the way he'd
have known a dragon if he saw one: they matched the
description of countless fantasy novels in most metrics, if
not all. Far from green skinned, the short humanoids were
almost as pale as the marble behind them, but they had
elongated, pointed nose and ears, extremely long, dexterous
looking fingers, and slanted, piercing eyes.

Harry tried not to stare as he and Professor McGonagall
approached the stairs leading up to the door. It took all
his self-control to keep his questions to himself.
Here were living beings, apparently just as sapient as
humans, but clearly from a drastically different genealogy!
He wondered how much their DNA resembled humans,
and if the two species were close enough genetically to
interbreed. A handful of goblin bones would probably drive
Richard Dawkins into an academic frenzy, let alone sight of
the real thing.

Above the great double doors was a gold and mahogany
plaque bearing the symbol of an ornate key, the name
\emph{Gringotts} inscribed on it. Below that were the words
\emph{Fortius Quo Fidelius}. Harry consulted his scattered Latin—
Stronger Through Loyalty?

"Good morning," he said to the goblins, who bowed in
response. The doors appeared to be thick, heavy marble,
but one of the goblins gripped its lower handle and swung
it open with ease, though he appeared no more muscular
than Harry. \emph{Mental note: size does not correlate with
strength in the magical world.}

Harry and Professor McGonagall walked through the doors
side by side, and the goblin closed the doors behind them. They
were in a small entrance hall that was mostly empty,
though curiously there were a pair of fireplaces to either
side. In front of them was another set of doors, these
silver, also flanked with goblins. As Harry approached, he
saw writing engraved on it:

\begin{inscription}
Enter, stranger, but take heed\\
Of what awaits the sin of greed\\
For those who take, but do not earn,\\
Must pay most dearly in their turn.\\
So if you seek beneath our floors\\
A treasure that was never yours,\\
Thief, you have been warned, beware\\
Of finding more than treasure there.
\end{inscription}

Harry swallowed. It should have seemed silly, something
out of a nursery tale{\el} but standing in what seemed
literally a goblin stronghold, the words conveyed a quiet,
self-assured threat that sent a shiver up Harry's spine.
\emph{Further note: do not antagonize a goblin.}

"Greetings, Madame McGonagall," said the goblin to their
right. His reedy voice bore the accent of a dialect
Harry had never heard before. "Greetings, Master Potter."

"Greetings," Harry replied. He glanced at Professor McGonagall curiously.

"I informed the bank that we would be arriving today, so
they would have your family vault key readily located," she
explained. "It hasn't been accessed in a decade."

"Ah. Well, thank you for keeping my inheritance safe for
me," Harry told the goblin. He felt rather awkward claiming
money from parents he had never met, but Harry knew he
would need it for school supplies that were sure to be
very expensive. How much would a genuine magic item go
for in the Muggle world? People already paid exorbitant
prices on claptrap that didn't even work; the tooth cleaning
potion he'd seen in a shop window would probably sell for
hundreds of pounds, maybe thousands. He hoped he
would be able to afford his wand and books, at least.

"Merely our duty, Master Potter," the goblin said,
bowing again. Was it Harry's imagination, or was there a
slightly mocking tone to the goblin's words?

The goblin opened the door, revealing a long hall filled with
goblins and wizards. The latter mostly stood in queues,
while the former walked about with an air of urgent
business or stood behind podiums and desks that placed
them a head above their clients. A rather obvious bit of
over-compensation, but Harry didn't blame them in the slightest.

All this Harry saw through what appeared to be a thin film
of water, falling gently from somewhere above the other
side of the doors and trickling into a narrow grate on the floor.

"It's called the Thief's Downfall," Professor McGonagall said,
seeing Harry's hesitation. "It washes away many forms of
magical disguise, and will ensure we are who we appear
to be. Your scar will become visible after we pass through,
but I'll obscure it again when we leave." She walked
through and turned to wait for him.

Harry took a breath and stepped through with shut eyes and tight
shoulders, anticipating a cold bath. However, the water felt entirely
lukewarm and evaporated immediately. Within seconds, he felt completely
dry, and he opened his eyes in astonishment.

Professor McGonagall gave him a brief smile, then led the
way to a podium labeled "Special Appointments." Harry
followed, trying not to stare at any one thing too long. He
saw a witch weighing a jingling pouch in her hand, goblins
writing on long sheets of parchment with feathered quills,
and a wizard handing an emerald the size of Harry's fist
to his goblin teller, who took out a monocle to examine the gem.

The goblin they approached seemed older than the others,
head mostly bald and hair long and white, with small
spectacles perched on his long nose. "Yes?" he asked,
without glancing up from the parchment he was examining.

"Harry Potter is here to access his vault."

There was a slight pause, and the goblin's eyes flicked
upward at Harry. "You have the key?"

Professor McGonagall pulled an iron key out of her sleeve
and held it over the desk.

The older goblin squinted at the key for a
moment, running his thin fingers over the engraving,
then he handed it back. "Very good." The goblin reached to
the side and picked up a bell, which he rang in a
deliberate pattern: \emph{ding-a-dong, ding-ding, dong-ding!}
"Griphook will guide you." He rolled his parchment up a bit
and continued reading. Harry saw that it extended all the
way down to the floor. He wondered why a society that
clearly had access to books would still use scrolls. Perhaps
it was a goblin thing, along with their aversion to fountain pens.

Griphook proved to be a youngish goblin, with relatively
smooth skin and full dark hair slicked back over his head.
He bowed after approaching them. "Madame McGonagall.
Master Potter. Follow me, please."

The goblin led them out of the main hall through a side
door, which opened into a downward staircase. The stairs
were clean white marble at first, but soon changed to dark
stone, and the glowing chandeliers gave way to flaming
torches. Harry knew they must have been enchanted, as
there was no smoke, which would have filled the tunnel
without some airway to escape through.

\emph{I'm walking down a goblin tunnel,} Harry thought with a
renewed touch of unreality. The tunnel soon leveled off,
and then the walls fell away to reveal long, twisting paths
with rail tracks along the floor. Mine carts sat on short
tracks that branched off the main lines, and Griphook led
them to one.

"Please step aboard, Master Potter."

Harry met the goblin's gaze. "Mister will suffice, thank
you." Harry remembered vacationing with his parents at an
expensive hotel once. They'd been waited on by the staff
with an overwhelming deference that was enjoyable, yet
occasionally uncomfortable. Though the thought of
having servants (or better yet, minions) was appealing on
a number of levels, he didn't quite feel comfortable being
called "master" by another race. Especially not coupled with
the vaguely mocking tone he sensed from the goblins,
though perhaps that was merely a difference in
mannerisms or accent that didn't quite translate well. "And
my full name is Potter-Evans-Verres."

The goblin peered back at him for a silent moment, then
another. Harry didn't drop his gaze, and the goblin finally
inclined his head briefly. "As you say, Mr.~Potter-Evans-Verres."

Harry followed Professor McGonagall onto the trolley,
ignoring her speculative look. He noted the lack of hand
rails or seat belts, and began to feel nervous as Griphook
stepped in and closed the side hatch. "Is this really the
safest way to conduct a banking transaction?"

"No," Griphook said with a wide grin, revealing sharp teeth.
He pulled a key out of his vest pocket, identical to the
one Professor McGonagall had, and inserted it into a
keyhole at the back of the trolley. "But then, it wouldn't
be much good if it were." He twisted the key, and the
mine cart shuddered, rolled slowly onto the main track,
then \emph{zoomed} forward at what only \emph{felt} like roughly half
the speed of sound.

Harry's yell of surprise was soon lost behind them as the
wind whipped it from his lips, and he gripped the sides of
the cart until his knuckles turned white. He gave Professor
McGonagall an accusatory glare, and the witch merely
raised an eyebrow, arms crossed nonchalantly over her
chest as her lips twitched at the corners. Harry grit his
teeth and slowly drew his hands back to his sides, ignoring
the lurching sensations in his stomach as the cart rocketed
through the twisting caverns, down, down, down.

\emph{Of course, they'd assure it's safe if you have the proper
means to travel it}, Harry chided himself. They passed by
glittering crystal outgrowths, various sized vault doors, and
through another Thief's Downfall, this one much bigger and
ice cold, though he once again dried in seconds.

Harry raised his voice over the clatter of the wheels. "Do
all wizards keep their money here?"

"All that care about their gold!" Griphook replied.

"Gold?"

"And silver, and bronze!" Professor McGonagall called out.
"The gold coins are called Galleons, the silver Sickles, and
the bronze Knuts. It's twenty-nine Knuts to a Sickle,
seventeen Sickles to a Galleon!"

Harry was still processing the implications of this as they
shot around a particularly wide corner, and his next
question was interrupted by a burst of fire that illuminated
the darkness around them. Harry twisted his head around,
but the chamber was already out of sight. "What was
that?" he yelled back at Griphook.

"Just a dragon!" Griphook said with a smirk, and Harry
couldn't tell if he was joking or not. His mind raced in a
whole new direction, monetary matters temporarily tabled.

Soon after, the cart began to slow, its deafening rattle
quieting little by little until it finally took one of the
branching paths. The trolley coasted along it and reached a
relatively small vault door with no markings on it. Griphook
hopped out of the cart and walked up to the door, then
inserted the same key he'd put into the trolley cart. A
second keyhole was parallel to it, and Professor McGonagall
inserted her, or rather Harry's, key into that one. They
twisted together, there was a heavy metal \emph{clunk,} and the
door swung inward.

Harry stepped inside the brightly lit marble room and felt
his jaw unhinge in a gape.

Heaps of gold Galleons. Stacks of silver Sickles. Piles of
bronze Knuts. More money than he'd ever seen in one
place, all in the form of gleaming treasure that would make
Bluebeard jealous. \emph{This{\el} is all mine?}

Harry was vaguely aware of Professor McGonagall leaning
casually against the wall, eyes intent. Watching him. Well, that made
sense. Being plopped in front of a giant heap of gold coins was a test of
character so pure it was archetypal.

Harry closed his mouth. First things first{\el} get an estimate
of how much money he was actually looking at, in a way
he could understand. "Are these coins the pure metal?" he asked Griphook.

"What?" the goblin spat from the doorway, voice harsh. "Are you
questioning the integrity of Gringotts, Mr.~Potter-Evans-Verres?"

"No," said Harry absently, "not at all, sorry if that came out wrong, sir. I
just have no idea at all how your financial system works. I'm asking if
Galleons in general are made of pure gold."

"Of course," said Griphook.

"And can anyone coin them, or are they issued by a monopoly that thereby
collects seigniorage?"

"What?" said Professor McGonagall.

Griphook grinned, showing sharp teeth. "Only a fool would trust any but goblin
coin!"

"In other words," Harry said, "the coins aren't supposed to be worth any more
than the metal making them up?"

Griphook stared at Harry. Professor McGonagall looked bemused.

"I mean, suppose I came in here with a ton of silver. Could I get a ton of
Sickles made from it?"

"For a fee, Mr.~Potter-Evans-Verres." The goblin watched him with glittering
eyes. "For a certain fee. Where would you find a ton of silver, I wonder?"

"I was speaking hypothetically," Harry said. \emph{For now, at any rate.}
"So{\el} how much would you charge in fees, as a fraction of the whole
weight?"

Griphook's eyes were intent. "I would have to consult my superiors{\el}"

"Give me a wild guess. I won't hold Gringotts to it."

"A twentieth part of the metal would well pay for the coining."

Harry nodded. "Thank you very much, Mr.~Griphook."

\emph{So not only is the wizarding economy almost completely decoupled from the
Muggle economy, no one here has ever heard of arbitrage.} The larger Muggle
economy had a fluctuating trading range of gold to silver, so every time the
Muggle gold-to-silver ratio got more than 5\% away from the weight of seventeen
Sickles to one Galleon, either gold or silver should have drained from the
wizarding economy until it became impossible to maintain the exchange rate.
Bring in a ton of silver, change to Sickles (and pay 5\%), change the Sickles
for Galleons, take the gold to the Muggle world, exchange it for more silver
than you started with, and repeat.

Wasn't the Muggle gold to silver ratio somewhere around fifty to one? Harry
didn't think it was seventeen, anyway. And it looked like the silver coins were
actually \emph{smaller} than the gold coins.

Then again, Harry was standing in a bank that \emph{literally} stored your
money in vaults full of gold coins guarded by dragons, where you had to go in
and take coins out of your vault whenever you wanted to spend money. The finer
points of arbitraging away market inefficiencies might well be lost on them.
He'd been tempted to make snide remarks about the crudity of their financial
system{\el}

\emph{But the sad thing is, their way might actually be better.}

On the other hand, one competent hedge fundie could probably own the whole
wizarding world within a week. Harry filed away this notion in case he ever ran
out of money, or had a week free.

Meanwhile, the giant heaps of gold coins within the Potter vault would suit
his near-term requirements.

Harry stepped forward and began picking up gold coins with one hand and
dumping them into the other.

When he had reached twenty, Professor McGonagall coughed. "I think that will be
more than enough to pay for your school supplies, Mr.~Potter."

"Hm?" Harry said, his mind elsewhere. "Hold on, I'm doing a Fermi calculation."

"A \emph{what?}" said Professor McGonagall, sounding somewhat alarmed.

"It's a mathematical thing. Named after Enrico Fermi. A way of getting rough
numbers quickly in your head{\el}"

Twenty gold Galleons weighed a tenth of a kilogram, maybe? And gold was, what,
ten thousand British pounds a kilogram? So a Galleon would be worth about fifty
pounds{\el} The mounds of gold coins looked to be about sixty coins high and
twenty coins wide in either dimension of the base, and a mound was pyramidal,
so it would be around one-third of the cube. Eight thousand Galleons per mound,
roughly, and there were around five mounds of that size, so forty thousand
Galleons or 2 million pounds sterling.

Harry smiled with a certain grim satisfaction. It was too bad that he
was right in the middle of discovering the amazing new world of magic, and
couldn't take time out to explore the amazing new world of being rich, which a
quick Fermi estimate said was roughly a billion times less interesting.

\emph{Still, that's the last time I ever mow a lawn for one lousy pound.}

Harry wheeled from the giant heap of money. "Pardon me for asking, Professor
McGonagall, but I understand that my parents were in their twenties when they
died. Is this a \emph{usual} amount of money for a young couple to have in
their vault, in the wizarding world?" If it was, a cup of tea probably cost
five thousand pounds. Rule one of economics: you can't eat money.

Professor McGonagall shook her head. "Your father was the last heir of an old
family, Mr.~Potter. It's also possible{\el}" The witch hesitated. "Some of
this money may be from bounties placed on You-Know-Who, payable to his ki—ah,
to whoever might defeat him. Or those bounties might not have been collected
yet. I am not sure."

"Interesting{\el}" Harry said slowly. "So some of this really is, in a
sense, mine. That is, earned by me. Sort of. Possibly. Even if I don't remember
the occasion." Harry's fingers tapped against his leg. "That makes me
feel less guilty about spending \emph{a very tiny fraction of it! Don't panic,
Professor McGonagall!}"

"Mr.~Potter! You are a minor, and as such, you will only be allowed to make
\emph{reasonable} withdrawals from—"

"I am \emph{all about} reasonable! I am totally on board with fiscal prudence
and impulse control! But I \emph{did} see some things on the way here which
would constitute \emph{sensible, grown-up} purchases{\el}"

Harry locked gazes with Professor McGonagall, engaging in a silent staring
contest.

"Like what?" Professor McGonagall said finally.

"Trunks whose insides hold more than their outsides?"

Professor McGonagall's face grew stern. "Those are \emph{very} expensive,
Mr.~Potter!"

"Yes, but—" Harry pleaded. "I'm sure that when I'm an adult I'll want one.
And I \emph{can} afford one. Logically, it would make just as much sense to buy
it now instead of later, and get the use of it right away. It's the same money
either way, right? I mean, I \emph{would} want a good one, with \emph{lots} of
room inside, good enough that I wouldn't have to just get a better one
later{\el}" Harry trailed off hopefully.

Professor McGonagall's gaze didn't waver. "And just what would you \emph{keep}
in a trunk like that, Mr.~Potter—"

"Books."

"Of course," sighed Professor McGonagall.

"You should have told me \emph{much earlier} that sort of magic item existed!
And that I could afford one! Now my father and I are going to have to spend the
next two days \emph{frantically} hitting up all the secondhand bookshops for
old textbooks, so I can have a decent science library with me at Hogwarts—and
maybe a small science fiction collection, if I can assemble something decent
out of the bargain bins. Or better yet, I'll make the deal a little sweeter for
you, okay? Just let me buy—"

"\emph{Mr.~Potter!} You think you can \emph{bribe} me?"

"What? \emph{No!} Not like that! I'm saying, Hogwarts can keep some of the
books I bring, if you think that any of them would make good additions to the
library. I'm going to be getting them cheap, and \emph{I} just want to have
them around somewhere or other. It's okay to bribe people with \emph{books,}
right? That's a—"

"Family tradition."

"Yes, exactly."

Professor McGonagall's body seemed to slump, the shoulders lowering within her
black robes. "I cannot deny the sense of your words, though I much wish I
could. I will allow you to withdraw an additional hundred Galleons,
Mr.~Potter." She sighed again. "I \emph{know} that I shall regret this, and I
am doing it anyway."

"That's the spirit! And does a `mokeskin pouch' do what I think it does?"

"It can't do as much as a trunk," the witch said with visible reluctance,
"but{\el} a mokeskin pouch with a Retrieval Charm and Undetectable Extension
Charm can hold a number of items until they are called forth by the one who
placed them—"

"Yes!" Harry's excitement made him shift from foot to foot,
eyes alight. "I definitely need one of those too! 
Batman's utility belt of holding! Never mind my Swiss
army knife, I could carry a whole tool set in there! Or \emph{books!} I could
have the top three books I was reading on me at all times, and just pull one
out anywhere! I'll never have to waste another minute of my life! What do you
say, Professor McGonagall? It's for the sake of children's reading, the best of
all possible causes."

"{\el}I suppose you may add another ten Galleons."

Griphook was favoring Harry with a gaze of frank respect, possibly even
outright admiration.

"And a little spending money, like you mentioned earlier. I think I can
remember seeing one or two other things I might want to store in that pouch."

"\emph{Don't push it, Mr.~Potter.}"

"But oh, Professor McGonagall, why rain on my parade? Surely this is a
\emph{happy} day, when I discover all things wizarding for the first time! Why
act the part of the grumpy grownup when instead you could smile and remember
your own innocent childhood, watching the look of delight upon my young face as
I buy a few toys using an insignificant fraction of the wealth that I earned by
defeating the most terrible wizard Britain has ever known, not that I'm
accusing you of being ungrateful or anything, but still, what are a few toys
compared to that?"

"\emph{You,}" growled Professor McGonagall. There was a look on her face so
fearsome and terrible that Harry squeaked and stepped back, knocking over a
pile of gold coins with a great jingling noise and sprawling backwards into a
heap of money. Griphook sighed and put a palm over his face. "I would be doing
a great service to wizarding Britain, Mr.~Potter, if I locked you in this vault
and left you here."

And they left without any more trouble.
